---
title: Range-wide Assessment of *Rana boylii*

# Use letters for affiliations
author:
  - name: Ryan Peek
    affiliation: a
  - name: Michael Miller
    affiliation: b
address:
  - code: a
    address: Center for Watershed Sciences, University of California, Davis, CA, 95616
  - code: b
    address: Department of Animal Science, University of California, Davis, CA, 95616

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Peek and Miller

# Place DOI URL or CRAN Package URL here
doi: "https://cran.r-project.org/package=YourPackage"

# Abstract
abstract: |
  As freshwater systems and biota are rapidly declining globally; conservation efforts will require assessment of the adaptive capacity of populations to rapid environmental change. Small populations with limited genetic diversity may have reduced adaptive potential and difficulty responding to future environmental change. Identifying these populations is crucial for effective conservation prioritization and management. We characterize population structure and genetic variation in *Rana boylii* across the geographically rich landscale of California in Oregon. We identify 5 major genetically distinct groups and provide further detail within each group using RADSeq RAPTURE with over 8,000 putative loci. 

# Acknowledgements
acknowledgements: |
  This work builds upon work by Amy Lind and Brad Shaffer and extends work from [Lind et al.](https://link.springer.com/article/10.1007/s10592-010-0138-0), and uses the excellent  [rticles](https://cran.r-project.org/package=rticles) and [pinp](http://dirk.eddelbuettel.com/code/pinp.html) packages.

# One or more keywords
keywords:
  - rivers
  - flow regime
  - dams
  - genomics
  - frogs

# Bibliography TODO(also still needed in body) # instead of pinp use ~/Dropbox/JabRef/Peek-bibkey-2016.bib
bibliography: ~/Dropbox/JabRef/Peek-bibkey-2016

# Enable a watermark on the document
watermark: true

# Customize footer, eg by referencing the vignette
footer_contents: "YourPackage Vignette"

# Our style
output: pinp::pinp
---

## Introduction {#intro .unnumbered}

This *pinp is not PNAS* template started when the introduction to
[Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) by \cite{PeerJ:Rcpp} 
was converted into this updated
[Rcpp Introduction](https://eddelbuettel.github.io/pinp/Rcpp-introduction.pdf)
vignette.  It is based on the
[pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template of the wonderful [rticles](https://cran.r-project.org/package=rticles) package by
\cite{CRAN:rticles}. The conversion from markdown to latex is facilitated by
[rmarkdown](https://cran.r-project.org/package=rmarkdown)
\citep{CRAN:rmarkdown} and [knitr](https://cran.r-project.org/package=knitr)
\citep{CRAN:knitr}. The underlying LaTeX macros are from
[pnas.org](http://www.pnas.org/site/authors/latex.xhtml).

The remainder of the document carries over from the corresponding
[pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template document. but has been edited and updated to our use case.  A
few specific tips follow.  In general, for fine-tuning some knowledge
of LaTeX is helpful.

Please note that LaTeX typesetting in two-column mode with several
floating objects can be fragile.  You may need to iterate.


## Author Affiliations {#author-affiliations .unnumbered}

Per common academic best practice, you can include your department,
institution, and complete address, with the ZIP/postal code, for each
author. Use lower case letters to match authors with institutions, as
shown in the example. Authors with an ORCID ID may supply this
information at submission.

## Document Options

We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermarking;

<!-- ## Format {#format .unnumbered}

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted. -->


## References {#references .unnumbered}

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `Chicago.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code {#inliner .unnumbered}

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures {#sec:figures .unnumbered}

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.8\textwidth, height=4in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) {#typeset .unnumbered}

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by tree backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typset directly simply by relying on pandoc.


## Single column equations {#single-column-equations .unnumbered}

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->


<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak 
