---
title: "03 PCA"
author: "Ryan Peek"
date: "*Updated: `r format(Sys.Date())`*"
output: 
  html_document:
    keep_md: true
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: cosmo # my favs are cerulean, flatly, spacelab, cosmo, lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

## Plot PCA

Following scripts 01 and 02, use 03 to plot PCA's interactively or statically.


### All Samples (n=5)

Rerun with equal samples (n=5), except TUO/STAN/PIT/CACHE.

```{r pcaAlln5, message=FALSE, warning=FALSE}

suppressPackageStartupMessages({library(tidyverse); library(here)})

# sftp in and cd projects/rangewide/pop_gen/results_pca
# get get ncoast_rabo*covMat

# source the function to plot
source(paste0(here(),"/scripts/functions/f_read_covar_rangewide.R"))

# load the metadata
metadat <- read_csv(file = paste0(here(), "/data/rapture_metadata_rabo.csv"))

# set site/reads for bamlist/covar filepaths:
reads <- "25k"
site <- "all_rabo_n4"
covarpath<- paste0(here(), "/data_output/pca/", site, "_", reads,".covMat")
bampath <- paste0(here(), "/data_output/bamlists/", site, "_", reads, "_thresh.bamlist")

# run function
(read_covar(covarpath, bampath, metadat, c(1,2), "ecoreg", "HUC6", plotlyplot = T))


# RAW GGPLOT

# read covar file
covar <- read.table(file = covarpath, stringsAsFactors = F)
# get bamlist for spp
bams <- read.table(bampath, stringsAsFactors = F, header = F) # read in the bamlist
bams$V2 <- sub('\\..*$', '', basename(bams$V1)) # remove the path and file extension
annot <- inner_join(bams, metadat, by=c("V2"="Seq")) %>% select(-V1) # join with the metadata
  
  # Eigenvalues
eig <- eigen(covar, symm=TRUE)
eig$val <- eig$val/sum(eig$val)
PC <- as.data.frame(eig$vectors)
colnames(PC) <- gsub("V", "PC", colnames(PC))
  
# these are all local variables that can be added...ultimately up to personal pref for plot
PC$Pop <- factor(annot$River)
PC$Locality <- factor(annot$Locality) 
PC$ID <- factor(annot$SampleID)
PC$HU_12_NAME <- factor(annot$HU_12_NAME)
PC$HU_8_NAME <- factor(annot$HU_8_NAME)
PC$HUC6 <- factor(annot$HUC_6)
PC$LabID <- factor(annot$LabID)
PC$ecoreg <- factor(annot$EcoRegion)
  
# set up plot  # PC's:
pc1 <- 1
pc2 <- 2
  
# Title: (% explained)
title <- paste("PC",pc1," (",signif(eig$val[pc1], digits=3)*100,"%)"," / PC", pc2," (",signif(eig$val[pc2], digits=3)*100,"%)", sep="",collapse="")

ggplot(data=PC, aes_string(x=paste0("PC",pc1), y=paste0("PC",pc2),
                           color=quote(ecoreg),
                           text=quote((paste0("ID: ", ID, " <br> River: ", Locality))))) +
  
  geom_point(size=4, alpha=0.8) +
  #ggforce::geom_mark_circle(aes(group=watershed))+
  theme_bw(base_family = "Roboto Condensed") +
  #theme(legend.position="bottom") +
  scale_color_viridis_d("River") + 
  ggtitle(paste0(title))

ggsave(filename = "figs/general_pca_rangewide_n5.png", width = 8, height=7, units = "in", dpi=300)

```


**Notes**

### North Coast

263 samples, not balanced.

```{r pcaNorthCoast, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))

# sftp in and cd projects/rangewide/pop_gen/results_pca
# get get ncoast_rabo*covMat

# source the function to plot
source(paste0(here(),"/scripts/functions/f_read_covar.R"))

# load the metadata
metadat <- read_csv(file = paste0(here(), "/data/rapture_metadata_rabo.csv"))

# set site/reads for bamlist/covar filepaths:
reads <- "25k"
site <- "ncoast_rabo"
covarpath<- paste0(here(), "/data_output/pca/", site, "_", reads,".covMat")
bampath <- paste0(here(), "/data_output/bamlists/", site, "_", reads, "_thresh.bamlist")

# run function
(read_covar(covarpath, bampath, metadat, c(1,3), "Pop", "HU_8_NAME", plotlyplot = TRUE))

```

A few notes:

 - The **SMITH-HurdyGurdy** seems to parse out differently from all other pops in the north coast, along with 
 - The **MAD-RedwoodCreek**
 - **RAP-109**: Gual: SF Gualala-Salmon, cluster with Mad and Trinity
 - **RAP-111**: On Mad river is different from other Mad samples on PCA.
 - **RAP-229**: Also seems to be an outlier compared to other points in SF Trinity Cluster
 - **RAP-321**: Seems to be an outlier that sits between the Smith and other pops
 - **RAP-344, RAP-335**: SF Eel Elder, seem a bit odd but cluster with Mad and Trinity samples
 - **RAP1418**: Also seems to be an outlier compared to other points within SMITH-HurdyGurdy Creek Cluster
 
 - The **Klamath** samples definitely cluster slightly closer to the SMITH samples than any of the other rivers (including MAD, MAT VanDUZ, SF Eel, Tinity)

### North Coast (n=4)

```{r pcaNorthCoastn4, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({library(tidyverse);
                               library(here)})

# sftp in and cd projects/rangewide/pop_gen/results_pca
# get get ncoast_rabo*covMat

# source the function to plot
source(paste0(here(),"/scripts/functions/f_read_covar_rangewide.R"))

# load the metadata
metadat <- read_csv(file = paste0(here(), "/data/rapture_metadata_rabo.csv"))

# set site/reads for bamlist/covar filepaths:
reads <- "25k"
site <- "ncoast_rabo_n4"
covarpath<- paste0(here(), "/data_output/pca/", site, "_", reads,".covMat")
bampath <- paste0(here(), "/data_output/bamlists/", site, "_", reads, "_thresh.bamlist")

# run function
(read_covar(covarpath, bampath, metadat, c(2,5), "Pop", "HU_8_NAME", plotlyplot = TRUE))

```

